### 2024-03-18 paper/paper.qmd
Echo false working now:)

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
execute:
  echo: FALSE 
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

//sum up gen survey and the questions the survey is intended to ask.
Background about the GSS.

Background about the sampling methodology

Background about the questions that are asked.

Plot each variable of interest and add discussion around them.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Results

Our results are summarized in.

//Add some analysis here. So that might be looking at difference over time or between states or between genders. Whatever.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# References

TESTING GROUND

```{r}
### Load required libraries
library(ggplot2)
library(tidyr)

```

```{r}

```

```{r,echo=FALSE, results=FALSE}
#| echo: false
#| warning: false
#| message: false

# Plot 1: GGPlot for GunLaw
Gunlaw <- read_csv(here("./data/analysis_data/Gunlaw.csv"))
transform(Gunlaw, 'Total Americans in Favour' = as.numeric('Total Americans in Favour'))

names(Gunlaw) <- c("Year", "Total")

ggplot(data=Gunlaw, aes(x = Year, y = Total, group = 1)) +
  geom_line()+
  labs(title = "Percentage of Americans in favour of police backed gun permits", y = "Total % of Americans in favour")

```